Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-05-23 paper-new/technical-background.tex
several fixes to paper

Ticked Cubical Type Theory \cite{mogelberg-veltri2019} is an extension of Cubical
Type Theory \cite{CohenCoquandHuberMortberg2017}
in which there is an additional sort called \emph{ticks}. Given a clock $k$,
a tick $t : \tick\, k$ serves as evidence that one unit of time has passed according to the clock $k$.
The type $\later A$ is represented as a function from ticks of a clock $k$ to $A$.
The type $A$ is allowed to depend on $t$, in which case we write $\later^k_t A$
to emphasize the dependence.

@@

The rules for tick abstraction and application are similar to those of dependent
$\Pi$ types. 
In particular, if we have a term $M$ of type $\later^k A$, and we
have available in the context a tick $t' : \tick\, k$, then we can apply the tick to $M$ to get
a term $M[t'] : A[t'/t]$. We will also write tick application as $M_t$.
Conversely, if in a context $\Gamma, t : \tick\, k$ we have that $M$ has type $A$,
then in context $\Gamma$ we have $\lambda t.M$ has type $\later A$. % TODO dependent version?

>>>>

Ticked Cubical Type Theory \cite{mogelberg-veltri2019} is an extension of
Cubical Type Theory \cite{CohenCoquandHuberMortberg2017} that has clocks as well
as an additional sort called \emph{ticks}. Ticks were originally introduced in
\cite{bahr-grathwohl-bugge-mogelberg2017}. Given a clock $k$, a tick $t :
\tick\, k$ serves as evidence that one unit of time has passed according to the
clock $k$. In Ticked Cubical Type Theory, the type $\later^k A$ is the type of
dependent functions from ticks of the clock $k$ to $A$. The type $A$ is allowed
to depend on $t$, in which case we write $\later^k_t A$ to emphasize the
dependence.

@@

The rules for tick abstraction and application are similar to those of ordinary
$\Pi$ types. In particular, if in a context $\Gamma, t : \tick\, k$ we have that $M$
has type $A$, then in context $\Gamma$ we have $\lambda t.M$ has type $\later^k
A$. % TODO dependent version?
%
Conversely, if we have a term $M$ of type $\later^k A$, and we have available in
the context a tick $t' : \tick\, k$, then we can apply the tick to $M$ to get a
term $M[t'] : A[t'/t]$. % TODO restriction on tick application
%
For the sake of brevity, we will also write tick application as $M_t$.
